%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Medium Length Professional CV
% LaTeX Template
% Version 2.0 (8/5/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Trey Hunner (http://www.treyhunner.com/)
%
% Important note:
% This template requires the resume.cls file to be in the same directory as the
% .tex file. The resume.cls file provides the resume style used for structuring the
% document.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass{resume} % Use the custom resume.cls style

\usepackage[left=0.75in,top=0.6in,right=0.75in,bottom=0.6in]{geometry}
\usepackage[none]{hyphenat}%%%%
% Document margins

\name{Peter Hennings} % Your name
\address{3905 Clawson Rd \\ Unit 6 \\ Austin, TX 78704} % Your address
% \address{123 Pleasant Lane \\ City, State 12345} % Your secondary addess (optional)
\address{(410)~$\cdot$~812~$\cdot$~1792 \\ phennings@gmail.com} % Your phone number and email

\begin{document}

%----------------------------------------------------------------------------------------
%	EDUCATION SECTION
%----------------------------------------------------------------------------------------

\begin{rSection}{Education}

{\bf Columbia University in the City of New York} \hfill {\em Did not finish} \\ 
Doctor of Philosophy Candidate, Computer Science \\
Pursued a PhD in Computer Science part-time via the CVN

{\bf The Johns Hopkins University} \hfill {\em May 2015} \\
Master of Science, Applied and Computational Mathematics

{\bf University of Delaware} \hfill {\em June 2008} \\
Bachelor of Science, Computer Science \\
Bachelor of Science, Mathematics \\

\end{rSection}

%----------------------------------------------------------------------------------------
%	WORK EXPERIENCE SECTION
%----------------------------------------------------------------------------------------

\begin{rSection}{Experience}

\begin{rSubsection}{TrailStone Group}{August 2018 - Present}{Quantitative
    Analyst}{Austin, TX}
  \item Model and structure bespoke power products for counterparties, with an
    emphasis on variable quantity risk (i.e. load, wind, solar).
  \item Empirical and fundamental modeling of short/long-term ERCOT nodal prices
    for the purposes of structuring hedges and proprietary financial trading.
\end{rSubsection}

\begin{rSubsection}{Constellation Energy}{June 2011 -
    August 2018}{Principal Quantitative Analyst}{Baltimore, MD}
\item Product owner of firm's risk system. Oversaw analytical
  overhaul to improve and unify the firm's
  distributional modeling of temperature, load and fuel/power price, amongst others.
\item Model and structure various physical and financial products
  designed to mitigate the firm's merchant risk.  Executed structures include
  load supply, quanto options,  unit-contingent power purchases (wind, solar,
  and fossil), and virtual/physical tolling agreements.
\item Collaborated with trading on quantitative optimization strategies to maximize value of the physical power generation and load portfolio.
\end{rSubsection}

%------------------------------------------------

%\begin{rSubsection}{Self Employed}{September 2007 - Present}{Algorithmic Trader}{Baltimore, MD}
%\item Systematically trade non-commodity futures, using my own capital via the
%  Interactive Brokers platform.
%\end{rSubsection}

%------------------------------------------------

\begin{rSubsection}{Constellation Energy}{August 2008 - May 2011}{Quantitative Developer}{Baltimore, MD}
\item Co-lead consolidation, rearchitecture, and development of fully automated
  pricing models for various retail load-following energy products.
\item Collaborated with the strategies (quant) group to analyze potential structural riders to
  existing products and analytical improvements to risk and pricing models.
\end{rSubsection}

%\begin{rSubsection}{Google}{May 2007 - September 2007}{Summer of Code Intern}{Mountain View, CA}
%\item Selected by Google to research, and develop a collaborative filtering recommendation system based on Bayesian Networks.
%\end{rSubsection}

\end{rSection}

%----------------------------------------------------------------------------------------
%	TECHNICAL STRENGTHS SECTION
%----------------------------------------------------------------------------------------

\begin{rSection}{Technical Strengths}

\begin{tabular}{ @{} >{\bfseries}l @{\hspace{6ex}} l }
Computer Languages & Python, Slang, MATLAB, C\#, Java \\
Databases & Oracle, SQL Server, sqlite, MongoDB \\
Tools & SecDb, git, emacs \\
Specialty Technologies & pandas, numpy, statsmodels, sklearn, cython \\
Energy Products & Weather Derivatives, Tolls, Variable Quantity Swaps
\end{tabular}

\end{rSection}

\end{document}
